Uncertainty on the galaxy-halo connection for Lyman-$\alpha$ emitters at $z=3.1$

### 2016-04-20 paper/ms.tex
conclusions

\begin{abstract}
We study the impact of cosmic variance and observational uncertainties
in constraining the mass and occupation fraction, \focc, of
dark matter halos hosting \ly\ Emitting  Galaxies (LAEs) at high redshift. 
We construct mock catalogs from an N-body simulation to match the 
typical size of observed fields at $z=3.1$ ($\sim 1 {\rm deg^2}$).
In our model a dark matter halo with mass in the range $\mmin
<M_{\mathrm h}<\mmax$ can only host one detectable LAEat most.    
We explore the parameter space determined by \mmin, \mmax\ and \focc\
with a Markov Chain Monte-Carlo algorithm using as observational
constraints the angular correlation function (ACF) and the LAEs number density. 
We find that the preferred minium and maximum masses in our model span
a wide range $10^{9.6}\hMsun\leq \mmin \leq 10^{11.0}\hMsun$ ,
$10^{10.9}\hMsun\leq \mmax \leq 10^{13.0}\hMsun$; followed by a
wide range in the occupation fraction $0.02\leq \focc \leq 0.30$.   
As a consequence the median mass, $M_{50}$, of all models consistent with
observations has a large uncertainty $M_{50} =
3.16^{+9.34}_{-2.37}\times 10^{10}$\hMsun, in agreement with the most recent
observational estimates.
However, we find that the same individual models have a relatively tight
mass gap, $\Delta M \equiv \log_{10}\mmax - \log_{10}\mmin$, of $\Delta
M\approx 0.3$dex.
We are also able show that \focc\ is uniquely determined by $M_{\rm
  min}$, regardless of $M_{\rm max}$, by
$\focc\sim0.1(\mmin/10^{10.5}\hMsun )^{0.93}$.  
However, the current observational data only allows us to put weak
constraints on \mmin\ and consequently on \focc.
We show that upcoming large surveys covering at least $50$ deg$^{2}$
will be able to put tighter constraints on \mmin\ and \focc\ through the LAE number density
distribution width, $W_{1\sigma}$, constructed over several fields of
$\sim 1$ deg$^{2}$ using the relationship, $W_{1\sigma} =
0.138\times (\mmin/10^{11}\hMsun )^{0.177}$, calibrated from our results.
\end{abstract}

@@

The parameter $\focc$ can be thought as the LAE occupation fraction of
dark matter halos.
An astrophysical interpretation of $f_{\rm occ}$ convolves at least
three phenomena: the actual presence of a star forming galaxy in a
halo, the escape fraction of \ly\ radiation and its detectability as
a LAE.   
We do not try to disentangle these complex phenomena.
We instead opt for an arithmetic interpretation by setting $\focc$ as
the ratio between the observed number of LAEs and the number of halos
within the considered mass range, that is $\focc \equiv N_{\rm
  LAEs}/N_{\mathrm{halos}}$.

For each mock catalog we also randomly remove $22\%$ of the mock-LAEs and replace 
them with randomly distrubuted point to mimic the
effect of contamination in \citet{Bielby16} observations.
On top of that we apply rejection sampling  to our LAE selection 
along the radial direction
taking the transmission function of the NB479 filter used in \citet{Bielby16}
observations.

@@

In this letter we studied the impact of cosmic variance and
observational uncertainties in constraining the mass range and
occupation fraction of dark matter halos hosting  LAEs.   
To this end, we used the BolshoiP N-body simulation to construct  27
mock fields  with the same typical size  of observed fields at
$z=3.1$ ($\sim 1 {\rm deg^2}$).   
In our model a dark matter halo with mass in the range $M_{\rm
  min}<M_{\mathrm h}<M_{\rm   max}$ can only host one detectable LAE
at most.  
We explored the parameter space determined by \mmin\ and \mmax\ using
Monte Carlo Markov-Chain minimization to match the observed  ACF and
mean number density of LAEs.

Our analysis allowed us to put weak constraints on $M_{\rm min}$,
$M_{\rm max}$ and $f_{\rm occ}$ where $10^{9.7}\hMsun\leq \log M_{\rm
  min}\leq 10^{11.2}\hMsun$, $10^{10.9}\hMsun\leq \log M_{\rm max}\leq
10^{13.0}\hMsun$ and $0.02\hMsun\leq f_{\rm occ}\leq 0.5$, spanning
three orders of magnitude in halo mass. 
Previous works\citep{Hayashino2004, Gawiser07,Ouchi2008,Bielby16} have
found typical masses within somehow narrower mass ranges ($\sim
10^{10.5}$-$10^{12}$).   
The main reason for our weaker constraints and large discrepancies
with previous works resides in the cosmic variance on the typical
field size in observations.  
A thorough exploration of cosmic variance impact as we present in this
letter had not been presented so far in the literature.

Our analysis also allowed us to draw three results that can be used to
put tighter constraints on $M_{\rm min}$, $M_{\rm max}$ and $f_{\rm
  occ}$ once upcoming large LAE surveys, as the HETDEX project
\citep{Hetdex2011} and the HSC ultra deep survey, are available : 
\begin{enumerate}
\item $f_{\rm occ}$ and the median mass of LAEs are almost uniquely
  determined by $M_{\rm min}$ regardless of $M_{\rm max}$. \item
  measuring the width of the LAE number distribution function obtained
  over several fields of $\approx 1$ deg$^2$ one will be able to
  tightly constrain  $M_{\rm min}$, $M_{\rm med}$ and $f_{\rm occ}$ up
  to a factor of  2.  
\item  $M_{\rm max}$ drives the ACF strength at short angular
  distances ($<1\,\rm arcmin$), therefore precise measurements of the
  ACF at such scales are crucial to accurately determine \mmax. 
\end{enumerate}

@@

>>>>

\begin{abstract}
We study the impact of cosmic variance and observational uncertainties
in constraining the mass and occupation fraction, \focc, of
dark matter halos hosting \ly\ Emitting  Galaxies (LAEs) at high redshift. 
We construct mock catalogs from an N-body simulation to match the 
typical size of observed fields at $z=3.1$ ($\sim 1 {\rm deg^2}$).
In our model a dark matter halo with mass in the range $\mmin
<M_{\mathrm h}<\mmax$ can only host one detectable LAEat most.    
We explore the parameter space determined by \mmin, \mmax\ and \focc\
with a Markov Chain Monte-Carlo algorithm using as observational
constraints the angular correlation function (ACF) and the LAEs number density. 
We find that the preferred minium and maximum masses in our model span
a wide range $10^{9.6}\hMsun\leq \mmin \leq 10^{11.0}\hMsun$ ,
$10^{10.9}\hMsun\leq \mmax \leq 10^{13.0}\hMsun$; followed by a
wide range in the occupation fraction $0.02\leq \focc \leq 0.30$.   
As a consequence the median mass, $M_{50}$, of all models consistent with
observations has a large uncertainty $M_{50} =
3.16^{+9.34}_{-2.37}\times 10^{10}$\hMsun, in agreement with the most recent
observational estimates.
However, we find that the same individual models have a relatively tight
mass gap, $\Delta M \equiv \log_{10}\mmax - \log_{10}\mmin$, of $\Delta
M\approx 0.3$dex.
We are also able show that \focc\ is uniquely determined by $M_{\rm
  min}$, regardless of $M_{\rm max}$, by
$\focc\sim0.1(\mmin/10^{10.5}\hMsun )^{0.93}$.  
However, the current observational data only allows us to put weak
constraints on \mmin\ and consequently on \focc.
We argue that upcoming large surveys covering at least $25$ deg$^{2}$
should be able to put tighter constraints on \mmin\ and \focc\ through the LAE number density
distribution width, $W_{1\sigma}$, constructed over several fields of
$\sim 1$ deg$^{2}$ using the relationship, $W_{1\sigma} =
0.138\times (\mmin/10^{11}\hMsun )^{0.177}$, calibrated from our results.
\end{abstract}

@@

An astrophysical interpretation of $f_{\rm occ}$ convolves at least
four phenomena: the actual presence of a star forming galaxy in a
halo, a duty cycle in the star formation rate, the escape fraction of
\ly\ radiation and its detectability as a LAE.   
We do not try to disentangle these effects.
We instead opt for a purely arithmetic interpretation by setting
$\focc$ as the ratio between the observed number of LAEs and the
number of halos within the considered mass range, that is $\focc \equiv N_{\rm
  LAEs}/N_{\mathrm{halos}}$.

For each mock catalog we also randomly remove $22\%$ of the mock-LAEs and replace 
them with randomly distrubuted points to mimic the effect of
contamination in \citet{Bielby16} observations. 
On top of that we apply rejection sampling  to our LAE selection 
along the radial direction taking the transmission function of the
NB479 filter used in \citet{Bielby16} observations.

@@

In this letter we studied the impact of cosmic variance and
observational uncertainties in constraining the mass range and
occupation fraction of dark matter halos hosting  LAEs.   
To this end we used the BolshoiP N-body simulation to construct  27 mock fields
with the same typical size  of observed fields at $z=3.1$ ($\sim 1
{\rm deg^2}$).      
In our model a dark matter halo with mass in the range $M_{\rm
  min}<M_{\mathrm h}<M_{\rm   max}$ can only host one detectable LAE
at most.  
We explored the parameter space determined by \mmin\ and \mmax\ using
Monte Carlo Markov-Chain minimization to match the observed  ACF and
mean number density of LAEs.    
It is the first time that such a thorough exploration of the cosmic
variance impact on LAEs statistics is presented in the literature.

We find that once cosmic variance is taken into account, the
observational data can only put weak constraints on \ \mmax\ and \focc. 
When we translate these loos constraints into the median mass, we find
a general consistency with previous works \citep{Hayashino2004,
  Gawiser07,Ouchi2008} specially with the most recent one
\citep{Bielby16}.

Nevertheless, our analysis allowed us to draw three results that can
be used to put tighter constraints on \mmin, \mmax and \focc once
upcoming large LAE surveys, such as the HETDEX project 
\citep{Hetdex2011} and the HSC ultra deep survey, are available: 
\begin{enumerate}
\item \focc is uniquely determined by \mmin, regardless of
  \mmax. A precise determination of \mmin\ will thus fix \focc.
\item The width of the LAE number distribution function
  obtained over several fields of $\approx 1$ deg$^2$ is tightly
  correlated with \mmin. That measurement with next generation surveys
  will be able to constrain \mmin within a factor of $\sim 2$.
\item  \mmax drives the ACF strength at short angular
  distances ($<1\,\rm arcmin$), therefore precise measurements of the
  ACF at such scales are crucial to accurately determine \mmax.  
\end{enumerate}

@@

JEFR acknowledges financial support from Vicerrector\'ia de
Investigaciones at Uniandes from a a FAPA project.
